---
title: "Incarceration Trends in the U.S."
author: "SAMIRA SHIRAZY"
date: "2/14/2022"
output: html_document
---

``` {r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r sourcing, include = FALSE}
source("../source/analysis.R")
```

## Overview

This is a report regarding incarceration data for each state in the United
States from 1970 to 2018. ([source: Vera Institute of Justice](https://www.vera.org/)). The data reveals that the average percent of Black people in jail since 1970 has been `r percent_black_jail`, whereas the percent of their white counterparts has been `r percent_white_jail`. In 2000, the highest incarceration rate for Black people was `r black_highest_2000` and `r white_highest_2000` for white people. From 1985 to 2016, the total jail population in the U.S. has changed by `r difference_1970_2018`. The highest jail rated capacity was `r highest_jrc_value` in `r highest_jrc_year`. This report will explore the trends of incarceration with regards to factors like: race, year, etc. 

### Incarceration by Race Time Trend
```{r time_trend, echo = FALSE, message = FALSE, warning = FALSE}
plot(time_trend_chart)
```

This chart displays the jail population distribution with regards to the major racial groups. The three that make up the most are: White, Black, and Latinx. Following the trend of these three show that the proportion that each of these racial groups make out of the total jail population has stay pretty consistent in this timeline. Even around the year 1992-1993, their peaks align with one another. 

### Jail Rated Capacity versus Jail Population
```{r comparison, echo = FALSE, message = FALSE, warning = FALSE}
plot(comparison_chart)
```

This scatter plot compares the jail rated capacity to the actual amount of inmates in the jail. Jail rated capacity refers to the number of beds the facility has to house all the inmates. The importance of this chart is to show the crowdedness in jails. As I examined the chart, I noticed that there were some jails that were severely overpopulated. For example, there was a jail with  a capacity of 15,000, but have a population of a little more than 25,000. That's a difference of 10,000 people. It makes me question, how bad are the living conditions there? This chart also can spark the conversation of morality within the jail system. Regardless of the crime committed, it is not right that some of these jails are over capacity. This could be dangerous for numerous reasons. For example: are there enough police staffed in proportion to the population of inmates? What about the spread of diseases, as highly populated but small spaces can be a breeding ground for that. Bad living conditions can worsen the mood of inmates, leading to more fights and outrages that could have been solved if the occupancy was less than the intended capacity. 

### Proportion of Black People in Jail in the U.S.
##### in the year 2018
```{r map, echo = FALSE, message = FALSE, warning = FALSE}
plot(map_chart)
```

This is a map of the United States and shows the proportion of jail inmates that are Black from 2018. It is important to have a visualization that shows the disparities between the counties. According to the U.S. census, the population of Black people in America was 14.7% ([U.S. Census](https://www.census.gov/programs-surveys/acs/news/updates/2018.html)). As we examine the map, we can see areas where the disparities take place. Despite being a little more than ten percent of the population, the south east area has a disproportionately high percent of Black people in jail. It is very purple, which represents percents anywhere from 40-100. A pattern that emerges is that the eastern side of the U.S. has a higher percent of Black people in jail compared to the west. More specifically, the south east. In all, this map is an eye opener in regards to how the population of the Black community is effected by the jail system.