---
title: "Incarceration Trends"
author: "SAMIRA SHIRAZY"
date: "2/14/2022"
output: html_document
---

``` {r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r, include = FALSE}
source("../source/analysis.R")
```


## Overview
This is a report regarding incarceration data for each state in the United
States from 1970 to 2018. 
([source: Vera Institute of Justice](https://www.vera.org/)). The data reveals 
that the average percent of Black people in jail since 1970 has been 
`r percent_black_jail`, whereas the percent of their white counterparts
has been `r percent_white_jail`. In 2000, the highest incarceration rate 
for Black people was `r black_highest_2000` and `r white_highest_2000` for white
people. From 1985 to 2016, the total jail population in the U.S. has changed
by `r difference_1970_2018`. The highest jail rated capacity was
`r highest_jrc`. 

### Time Trend Title
```{r, echo = FALSE}
plot(time_trend_chart)
```

### Comparison Chart Title
```{r, echo = FALSE}
plot(comparison_chart)
```

### Map Title
```{r, echo = FALSE}
plot(map_chart)
```
